Investigating Cross-Sectional Size Distribution of Randomly Distributed 3D Spheres

### 2020-05-24 Paper/main.tex
First draft completed

\begin{abstract}
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
  Text Text Text Text Text Text Text Text Text Text Text Text Text Text Text
\end{abstract}

@@

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/box3d.pdf}
\caption{$750$ spheres of radius 0.7 distributed randomly in a cube of length 20,
  such that no sphere is touching or overlapping another sphere. A randomly
  oriented plane passes through the box and ``cuts'' through the
  spheres.}\label{fig:3dplot_plane}
\end{center}
\end{figure}

@@

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/circles.pdf}
\caption{Planar cross-section constructed using mock data. Spheres cut by the
  plane appear as circles on the plane, with radius dependent on the relative
  position of the spheres to the plane. This distribution of circles can be
  modelled.}\label{fig:circles}
\end{center}
\end{figure}

@@

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_01.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.1$
  each. There are relatively few points since the spheres are small, so there is
  relatively small chance that they will be cut by the plane. So only a few
  circles are visible on the plane. The theoretical distribution is shown, with
  the experimental data matching this distribution, showing a slow increase in
  number or circles as radius increases.}\label{fig:size1}
\end{center}
\end{figure}

@@

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_03.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.3$
  each. Compared to Figure~\ref{fig:size1}, there are many more experimental data
  points, since it is more likely that the plane will intersect these larger
  spheres. The experimental data more clearly resembles the theoretical
  distribution, showing a strong increase in number of circles as radius
  increases towards the maximum of 0.3.}\label{fig:size3}
\end{center}
\end{figure}

@@

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.7$
  each. We do not observe much difference in how well the experimental data
  resembles the theoretical distribution between this figure and
  Figure~\ref{fig:size3}. With larger sized spheres, the simulation becomes harder
  to run as it is more difficult to randomly place spheres in a finite space
  without overlap.}\label{fig:size7}
\end{center}
\end{figure}

\begin{figure}[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07_random.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $R$
  where $0.65 < R < 0.75$. The theoretical distribution is now different,
  although the experimental data still resembles this new distribution, showing
  a maximum around $R = 0.7$ and a sharp decrease after this. This is likely
  because there are more spheres with radii such that an overlap between $0$ and
  $0.7$ is possible, and relatively few spheres with radii such that a larger
  overlap is possible.}
\label{fig:random}
\end{center}
\end{figure}

% \begin{figure}[H]%[!ht]
% \begin{center}
% \includegraphics[width=0.45\textwidth]{./../Figures/750_07_NoNoise.pdf}
% \caption{Illustrations, graphs, and photographs may fit across both columns, if
%   necessary. Your artwork must be in place in the article.}
% \label{fig:nonoise}
% \end{center}
% \end{figure}

% \begin{figure}[H]%[!ht]
% \begin{center}
% \includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_100.pdf}
% \caption{Illustrations, graphs, and photographs may fit across both columns, if
%   necessary. Your artwork must be in place in the article.}
% \label{fig:2noise}
% \end{center}
% \end{figure}

% \begin{figure}[H]%[!ht]
% \begin{center}
% \includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_10.pdf}
% \caption{Illustrations, graphs, and photographs may fit across both columns, if
%   necessary. Your artwork must be in place in the article.}
% \label{fig:1noise}
% \end{center}
% \end{figure}

\begin{figure*}[ht!]
  \centering
  \subfloat[Noise 2 orders of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_100.pdf}
  \label{fig:2noise}}
  \centering
  \subfloat[Noise 1 order of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_10.pdf}
  \label{fig:1noise}}
  \caption{Distribution of circle radii on plane for $750$ balls of radius $0.7$
    each. Each recorded radius has a random noise added to it in order to
    determine how sensitive the data is to experimental
    error. Figure~\ref{fig:2noise} shows the distribution for an added noise 2
    orders of magnitude smaller than the recorded radius, while
    Figure~\ref{fig:1noise} is the same for noise 1 order of magnitude smaller. As
    we can see, the experimental data very resembles the theoretical
    distribution when noise of 2 orders of magnitude smaller are added. But this
    is not true when noise of 1 order of magnitude smaller is added, implying
    the data is very sensitive to experimental data.}
  \label{fig:noiseplots}
\end{figure*}

\section{Conclusions}
Text

>>>>

\begin{abstract}
  In this paper we present a 3-dimensional numerical simulation demonstrating
  that it is possible to reproduce the known planar cross-section distributions
  for Wicksell's Corpuscle Problem, thereby recovering 3D particle size
  distributions from 2D cross-sections. We investigate the effects of different
  particle sizes on the experimental data, and show this method is accurate for
  2 different starting 3D size distributions. We also present our findings on
  how sensitive this method is to experimental data and discuss its efficacy in
  a real world application such as identifying a material using a planar
  cross-section image from a microscope.
\end{abstract}

@@

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/box3d.pdf}
\caption{$750$ spheres of radius 0.7 distributed randomly in a cube of length 20,
  such that no sphere is touching or overlapping another sphere. A randomly
  oriented plane passes through the box and ``cuts'' through the
  spheres.}\label{fig:3dplot_plane}
\end{center}
\end{figure}

@@

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/circles.pdf}
\caption{Planar cross-section constructed using mock data. Spheres cut by the
  plane appear as circles on the plane, with radius dependent on the relative
  position of the spheres to the plane. This distribution of circles can be
  modelled.}\label{fig:circles}
\end{center}
\end{figure}

@@

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_01.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.1$
  each. There are relatively few points since the spheres are small, so there is
  relatively small chance that they will be cut by the plane. So only a few
  circles are visible on the plane. The theoretical distribution is shown, with
  the experimental data matching this distribution, showing a slow increase in
  number or circles as radius increases.}\label{fig:size1}
\end{center}
\end{figure}

@@

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_03.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.3$
  each. Compared to Figure~\ref{fig:size1}, there are many more experimental data
  points, since it is more likely that the plane will intersect these larger
  spheres. The experimental data more clearly resembles the theoretical
  distribution, showing a strong increase in number of circles as radius
  increases towards the maximum of 0.3.}\label{fig:size3}
\end{center}
\end{figure}

@@

The sharp rise in the theoretical distribution close to the maximum radius is
also clear in the experimental data. It can be understood intuitively that we
expect to see more cross-sections with radius close to radius of the spheres
since there is a greater probability of intersection with the centre of a sphere
than any other point on the sphere.

Increasing the size of the spheres further to a radius of $0.7$, we might
initially expect the experimental data to resemble the theoretical distribution
even more closely, but this is not what we see. Figure~\ref{fig:size7} presents
the planar distribution for $750$ spheres of radius $0.7$ each.

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $0.7$
  each. We do not observe much difference in how well the experimental data
  resembles the theoretical distribution between this figure and
  Figure~\ref{fig:size3}. With larger sized spheres, the simulation becomes harder
  to run as it is more difficult to randomly place spheres in a finite space
  without overlap.}\label{fig:size7}
\end{center}
\end{figure}

In Figure~\ref{fig:size7}, we see a similar experimental distribution to that of
Figure~\ref{fig:size3}, as opposed to a distribution that more closely resembles
the theoretical distribution. While initially this may seem contrary to the
notion that increasing size of spheres would increase the number of
intersections and thus give a better distribution, this may simply be due to the
fact that increasing the size of the spheres beyond a certain radius does not
provide continually better results. We observe a similar number of intersections
as the case with spheres of radius $0.3$, and thus it is possible to conclude
that the best experimental distribution is found for a spherical size that
ensures a significant number of intersections. Though continually increasing
spherical size provides diminishing returns.

Indeed, we found that making the sphere sizes much larger made the placing of
the spheres in a finite space more difficult, and exponentially increased the
time complexity of the placing algorithm. This is because if the placed spheres
are too large, while it is initially easy to place spheres, as more are placed,
more coordinates must be randomly drawn before one if found for the next sphere
to be placed into. It is also possible that the algorithm never finds a possible
coordinate for later spheres. One possible solution to this may be to use a
better placing algorithm. Choosing a tight packing algorithm \todo{cite} would
maximise the number of spheres of a given radius that can be placed. However,
this would likely result in a less than random distribution of spheres if the
spheres are all of identical sizes. This would likely cause some systematic
error in the data, and so it was not investigated here. If we randomised the
sizes of the spheres as we do in the next part of the investigation, the problem
become significantly more complicated. Tight packing algorithms for irregular
shaped objects are very difficult to create and computationally expensive
\todo{cite}.

We next randomised the radius of each sphere to be in the range $0.65 < R <
0.75$. The experimental data for this simulation is presented in
Figure~\ref{fig:random}.

\begin{figure}%[H]%[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth]{./../Figures/750_07_random.pdf}
\caption{Distribution of circle radii on plane for $750$ spheres of radius $R$
  where $0.65 < R < 0.75$. The theoretical distribution is now different,
  although the experimental data still resembles this new distribution, showing
  a maximum around $R = 0.7$ and a sharp decrease after this. This is likely
  because there are more spheres with radii such that an overlap between $0$ and
  $0.7$ is possible, and relatively few spheres with radii such that a larger
  overlap is possible.}
\label{fig:random}
\end{center}
\end{figure}

Figure~\ref{fig:random} presents the experimental data for 750 spheres with each
sphere having a radius $0.65 < R < 0.75$. As we can see, the theoretical
distribution has a very different shape to the case of constant sphere
radius. We see that the theoretical distribution has a sharp peak centred around
(but not precisely at) $0.65$, the lower limit of the uniform range the radii
were drawn from, and a sharp decrease after this point. This is likely due to
the higher number of spheres available in the box which could be intersected so
as to show a planar cross-section with radius $\approx 0.65$. In a range of
radii between $0.65 < R < 0.75$, all the spheres in the box can be intersected
so as to produce a circle of radius $0.65$, but there will be fewer spheres
which can be intersected to produce a circle of radius greater than this. So a
peak at this point and then a sharp decrease is expected.

The experimental data resembles the theoretical distribution, clearly showing a
peak and sharp decrease after. However, we also see that there is significant
variance between bins, and there is a large scatter around the theoretical
data. As explained for Figure~\ref{fig:size3}, it would be possible to reduce
this fluctuation by using fewer bins, at the cost of having fewer data
points. With a more powerful computer, or if the simulation is optimised and
written with a compiled language, there is also scope for significantly
increasing the size and number of spheres used, which would provide better
experimental data.

While there is significant fluctuation and scatter in all of the simulations
presented so far, the shapes of the distributions were mostly clear, and they
resembled the theoretical distribution. However, if the system is very sensitive
to experimental error, it may be difficult to identify distribution shapes, and
thus prevent us from properly identifying the starting distribution of spherical
radii. We therefore introduced some additional random error into the data. After
the radii of the circles on the plane are calculated, a small random number is
added to each radius. The random number is drawn from a uniform distribution in
the range $[\frac{-r}{n}, \frac{r}{n}]$, where r is the circle's radius and n is
the number of orders of magnitude we wish to choose the errors. We tested the
system with errors 2 orders of magnitude smaller than the radii and 1 order of
magnitude smaller than the radii. This is presented in
Figure~\ref{fig:noiseplots}.

\begin{figure*}%[ht!]
  \centering
  \subfloat[Noise 2 orders of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_100.pdf}
  \label{fig:2noise}}
  \centering
  \subfloat[Noise 1 order of magnitude smaller than radius]
  {\includegraphics[width=0.45\textwidth]{./../Figures/750_07_Noise_10.pdf}
  \label{fig:1noise}}
  \caption{Distribution of circle radii on plane for $750$ spheres of radius $0.7$
    each. Each recorded radius has a random noise added to it in order to
    determine how sensitive the data is to experimental
    error. Figure~\ref{fig:2noise} shows the distribution for an added noise 2
    orders of magnitude smaller than the recorded radius, while
    Figure~\ref{fig:1noise} is the same for noise 1 order of magnitude smaller. As
    we can see, the experimental data very resembles the theoretical
    distribution when noise of 2 orders of magnitude smaller are added. But this
    is not true when noise of 1 order of magnitude smaller is added, implying
    the data is very sensitive to experimental data.}
  \label{fig:noiseplots}
\end{figure*}

Figure~\ref{fig:2noise} presents the distribution for $750$ spheres or radius
$0.7$ each, with an added ``noise'' that is 2 orders of magnitude smaller than
each measured radius. We can see that the experimental data looks very similar
to the data presented in Figure~\ref{fig:size7}. The shape of the distribution
is also very clear and resembles the theoretical distribution.

Figure~\ref{fig:1noise} presents the same simulation but with an added noise
that is 1 order of magnitude smaller than each measured radius. Here we can see
that the experimental data is unrecognisable, and the shape of the distribution
is no longer clear. We thus conclude that the system is very sensitive to noise
greater than 2 orders of magnitude compared to the measured data.

However, it would be more prudent to use a Gaussian distribution to draw the
values of the noise. This is because it is likely that the many sources of error
possible in an experimental measurement of particle cross-sections in a material
will be convoluted together into a Gaussian distribution of errors due to the
central limit theorem \todo{cite}.

\section{Conclusions}
In this investigation we showed that it is possible to reproduce expected planar
distributions for the Corpuscle Problem with a 3-dimensional numerical
simulation of spheres distributed randomly in space, where a plane oriented
randomly intersects the spheres. We investigated how effective this method of
simulation is for different particle sizes, and showed that while initially
larger particles provide better data since they allow for more intersections,
continually increasing particle size becomes ineffective and computationally
expensive.

We showed it was possible to reproduce the expected planar distributions for two
starting spherical radii distributions: a Dirac delta distribution, in which all
the spheres have the same radius, and a bounded uniform distribution, in which
all the spheres' radii are drawn randomly from a uniform distribution. Both
produce different shaped planar distributions but both provide data that
resembles the theoretical distributions.

We also investigated how sensitive this method is to experimental error. We
introduced some random noise into the system from a random uniform distribution
and demonstrated that the system is very sensitive to noise greater than two
orders of magnitude relative to measured data. Thus it is important for any
experimentalist making measurements of particle cross-sections in a material to
be very careful with error handling, as any error larger than 2 orders of
magnitude compared to measured data will result in great difficulty in
reproducing the spherical radii distributions.

We therefore showed in this investigation that it is possible to use the known
solutions to Wicksell's Corpuscle Problem to better identify and classify
materials by using planar measurements to gain information about 3D particle
distributions.
